Lesson 5
========================================================


```{r Third Qualitative Variable}
pf <- read.delim('../data/pseudo_facebook.tsv')

suppressMessages(library(dplyr))

pf.fc_by_age_gender <- pf %>%
  group_by(age, gender) %>%  # Group by age and gender
  summarise(
    mean_friend_count = mean(friend_count),
    median_friend_count = median(friend_count),
    n = n()
  ) %>%
  ungroup()  # Remove grouping
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}

ggplot(pf.fc_by_age_gender, aes(x = age, y = median_friend_count, color = gender)) +
  geom_line(stat = "summary", fun.y = "median") +
  labs(
    x = "Age",
    y = "Median Friend Count",
    title = "Median Friend Count by Age and Gender"
  )
```


### Reshaping Data


```{r}
install.packages('reshape2')
library(reshape2)
library(dplyr)
library(tidyr)

pf.fc_by_age_gender.wide <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
library(ggplot2)

# Create the plot
ggplot(pf.fc_by_age_gender.wide, aes(x = age, y = female / male)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(
    x = "Age",
    y = "Female-to-Male Median Friend Count Ratio",
    title = "Ratio of Female to Male Median Friend Counts"
  )
```

***

### Third Quantitative Variable

```{r Third Quantitative Variable}
# konvertnu 'tenure' na date
pf$tenure <- as.Date(pf$tenure, origin = "1970-01-01")

# vypocitam 'year_joined' 
pf$year_joined <- as.integer(format(pf$tenure, "%Y"))
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
bucket_boundaries <- c(2004, 2009, 2011, 2012, 2014)

pf$year_joined.bucket <- cut(2014 - pf$tenure / 365, breaks = bucket_boundaries, 
                              labels = c("(2004, 2009]", "(2009, 2011]", "(2011, 2012]", "(2012, 2014]"))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
filtered_data <- subset(pf, !is.na(year_joined.bucket))

ggplot(filtered_data, aes(x = age, y = friend_count, color = year_joined.bucket)) +
  geom_line(stat = "summary", fun.y = median) +
  labs(
    x = "Age",
    y = "Median Friend Count",
    title = "Median Friend Count by Age and Year Joined Bucket"
  )
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
filtered_data <- subset(pf, !is.na(year_joined.bucket))

grand_mean <- mean(filtered_data$friend_count, na.rm = TRUE)

ggplot(filtered_data, aes(x = age, y = friend_count, color = year_joined.bucket)) +
  geom_line(stat = "summary", fun.y = "median") + 
  geom_line(aes(x = age, y = grand_mean), linetype = 2, color = "black") + 
  labs(
    x = "Age",
    y = "Median Friend Count",
    title = "Median Friend Count by Age and Year Joined Bucket"
  )
```



### Friendships Initiated
```{r Friendships Initiated}
filtered_data <- subset(pf, tenure >= 1)

filtered_data$mean_friendships_per_day <- filtered_data$friendships_initiated / filtered_data$tenure

ggplot(filtered_data, aes(x = tenure, y = mean_friendships_per_day, color = year_joined.bucket)) +
  geom_line(stat = "summary", fun.y = "mean") +
  labs(
    x = "Tenure (days)",
    y = "Mean Friendships Initiated per Day",
    title = "Mean Friendships Initiated per Day vs. Tenure"
  )

# geom smooth
ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket), method = "auto") +
  labs(
    x = "Tenure (days)",
    y = "Mean Friendships Initiated per Day",
    title = "Mean Friendships Initiated per Day vs. Tenure (Smoothed)"
  )
```

***

### Bias-Variance Tradeoff Revisited
```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```


```{r Histograms Revisited}
library(ggplot2)
yo <- read.csv('../data/yogurt.csv')

yo$id <- factor(yo$id)

ggplot(yo, aes(x = price)) +
  geom_histogram(binwidth = 10, fill = "blue", color = "black") +
  labs(
    x = "Yogurt Price",
    y = "Frequency",
    title = "Histogram of Yogurt Prices"
  )

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
library(dplyr)

all_purchases <- aggregate(strawberry ~ id, data = yo, sum)

all_purchases <- rename(all_purchases, all.purchases = strawberry)

yo <- merge(yo, all_purchases, by = "id", all.x = TRUE)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(yo, aes(x = time, y = price)) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.3, size = 2) +
  labs(
    x = "Time",
    y = "Price",
    title = "Scatterplot of Price vs. Time"
  )

```

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) + facet_wrap( ~ id) + geom_line() + geom_point(aes(size = all_purchases), pch = 1)
```

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

